The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-02-01 src/ms.tex
Mention times in CTA section
Apply suggestions from code review
Co-authored-by: P. L. Lim <2090236+pllim@users.noreply.github.com>

Briefly summarize new procedures and governance structure, new CoCo, election
process overview, etc.

@@

\secauthor{Lía Corrales, David Shupe, Kelle Cruz + Learn team}

\subsubsection{Current status and scope}

{\it Learn Astropy} is an umbrella term that acknowledges the broad educational efforts made by the Astropy Project, which are led by the Learn Team.
The efforts focus on developing online content and workshops covering astro-specific coding tasks in Python.
As described in \textsl{[Astropy-ii paper]}, There are four different types of Learn Astropy content: \textit{tutorials}, consisting of Jupyter Notebook lessons that are published in HTML format online; \textit{guides}, which are a series of lessons providing a foundational resource for performing a certain type of astronomical analyses; \textit{examples}, which are snippets of code that showcase a short task that can be performed with Astropy or an affiliated package; and \textit{documentation}, which is contained within the code base.
This categorization drives content development, infrastructure choices, and the appearance of the Learn Astropy website.
The Learn Team meets weekly to work on creating, expanding, improving these educational resources.

\begin{itemize}

\item {\bf Learn Astropy website:} The Learn Team re-launched the website in 2021 with a new infrastructure platform, built around full-text search and interactive filtering functionality, with the goal of making content discoverable as the Learn Astropy content catalog expands.
This work has been supported in part by the Dunlap grant.
We have adopted Algolia, a search-as-a-service cloud platform, to store the full-text and metadata records of Learn Astropy's content.
The new Learn Astropy website is a JavaScript (Gatsby/React) application that uses the Algolia service to power its search and filtering user interface.
Our Python-based application, Learn Astropy Librarian, populates data into the Algolia service. We tuned the Librarian around specific content formats (such as Jupyter Notebook-based tutorial pages and Jupyter Book-based guides) to more accurately index content and heuristically extract metadata.
A consequence of the new platform is that we now maintain and compile content separately from the website application itself, enabling new content types.
Tutorials, which as Jupyter Notebook-based, are now compiled into their own Learn Astropy sub-site using nbcollection.
Guides, which are Jupyter Book-based, are also deployed as separate websites using GitHub pages. This architecture opens future possibilities of indexing third-party content, hosted elsewhere, such as on institutional websites.

\item \textbf{Tutorials} Currently, we have \# of tutorials, spanning a wide range of astronomical topics and common tasks.

\item \textbf{Guide} {\it ccdproc} Jupyter book has been completed.

\item {\bf Workshops:} the Project has been conducting workshops at winter meetings
of the American Astronomical Society since AAS 225 in January 2015. Up to the
start of the coronavirus pandemic, these were full-day in-person workshops with
as many as 90 participants and a dozen facilitators from the project. During the
pandemic, these workshops were moved to an online format and split into
basic and advanced sessions. Additionally, beginning with the AAS 238 online
meeting, the workshops have been expanded to Summer AAS meetings. The Learn
team finds that the workshop audience is best found as AAS meetings as opposed
to more general Python meetings.

The Astropy Project provided another mode of community engagement at AAS Meetings 235 and 237 by organizing a NumFOCUS Sponsored Projects booth in the AAS Exhibit Hall. Funding for the exhibit hall was provided alternately by NumFOCUS and later by the Moore Foundation funding (\textbf{check this}). The booth hosted a series of Q\&A special sessions during AAS 235 and webinars during the virtual AAS 237 meetings, to provide the general astronomy community information and access to experts on a variety of open source astronomical tools.
\end{itemize}

\subsubsection{Learn vision for the future}

The Learn plan going forward is to continue to improve the tutorial website and facilate new content;
to index third-party tutorials; and to look for supportable opportunities to expand the reach of
Astropy workshops beyond the American Astronomical Society meetings.

@@

\section{Community Engagement}

\item {\bf User forums:} The Astropy Project has historically maintained a number of avenues for users to seek help or gain access to the developer community. This includes the Astropy mailing list, the Python in Astronomy Facebook group, and the Astropy Slack workspace. The Moore Foundation grant allows this Slack space to
be on a paid plan; additionally NumFOCUS has negotiated a special rate for
open-source projets. While the Slack and Astropy-dev mailing lists are primarily used to discuss the project direction and updates, it was noted that the use of a Facebook community could present a barrier to open source. The Astropy Project identified a need for a public, archived, searchable, and easily-to-navigate interface for users to ask for help Accordingly, we have commissioned
a Discourse site which is more open than the Facebook group and more user-oriented
than Astropy Slack. A benefit of the Moore Foundation grant is that Astropy
developers are able to invoice as independent contractors the time they
spend helping users on these forums.

>>>>

%Briefly summarize new procedures and governance structure, new CoCo, election
%process overview, etc.

As part of the process of developing Astropy into a long-term sustainable
product, and to improve transparency and accountability, the Project agreed to
write down and formalize our governance structure (partly supported by explicit
funding for this purpose - see \S \ref{sec:project-funding}). At the 2019
Astropy Coordination Meeting, input was gathered from participants on what governance
structures existed in the associated Open Source Software communities, and what
would fit well with the needs of Astropy. This led into a ``retreat'' planned
for March 2020, but due to the COVID-19 pandemic, this became a series of
virtual meetings of the ``Astropy Governance Working Group''.  This group
drafted the APE 0 document \citep{ape0}, which was then eventually ratified and
implemented by the ``Astropy Governance Implementation Working Group'' in Fall
2021. While the process emphasized flexibility and the ability to adapt to
changing circumstances, it is expected that this is the framework Astropy's
governance will operate in for at least the medium-term future.

The APE0 \citep{ape0} document lays out the principles of this governance
structure, so we refer the reader to that document for a more thorough
description.  However here we highlight some key elements. While many of these
principles were already de facto true or have been discussed organically (and
have been discussed in earlier papers in this series), the APE0-based governance
aimed to provide a single place where the community can agree as a starting
point. With this in mind, it highlights the developer and user community of the
Astropy Project as the ultimate sources of authority, as well as the core
principle of ``do-ocracy'' that those who do work for the Project (be it coding,
training, or other less concrete contributions) gain more influence on the
outputs of the Project by virtue of their effort.  APE0 adds, however, the
concept of ``voting members'' - a self-governed part of that community who are
entrusted to elect the Coordination Committee. While this committee has existed
from the inception of the project, APE0 establishes a formal voting process for
this committee, and explicitly outlines the rights and responsibilities of the
Coordination Committee. This role is mainly to facilitate consensus and act as
the decision maker when other mechanisms have failed.  However, it also includes
powers that either require central authority or secrets (e.g., passwords), but
AOE0 also charges the Committee to devolve responsibilities and seek community
input on these items as often as possible.

The first Coordination Committee election under these rules took place in Fall
2021, electing a mix of prior existing and new coordination committee members,
and was contested in the sense of more candidates than available slots. This
suggests the process is already working to serve the long-term interests of the
committee to both spread the coordination effort, and to ensure it is not
dominated by the same people for as long as the Project continues. While other,
more fine-grained governance improvements are planned for the future, it is
clear the foundation is now in place.

@@

\secauthor{Lía Corrales, David Shupe + Learn team}

\begin{itemize}
\item Current status and scope
\item Vision for the future (review what we said in v2.0 paper)
\item User forums and engagement with user base
\item Workshops
\end{itemize}

@@
